Improving the Replicability of Psychological Science Through Pedagogy

### 2016-09-23 paper/paper.Rmd

We eventually used a hybrid, case-by-case selection criterion that attempted to maximize the success of the project while staying within the constraints of our budget. For all criteria, experimenters recruited an additional 5% of participants to ensure we still achieved the desired criterion if some participants skipped through or did not complete the experiment.
^[We initially were allocated \$1,500 but one student contributed personal research funds, leading to a total cost of $\approx$ \$1,700 for all studies. We attempted to set payment for our studies at approximately \$6/hour based on timing estimated from Pilot A sessions.] 
In four cases, we powered the replication attempt to 80\% post-hoc power. In four
cases, we used the original sample size. In one case, the original experiment was extremely over-powered in the post-hoc analysis, so we chose a smaller sample size that still would yield high power. In two others, we reduced the number of trials as part of the adaptation to Mechanical Turk (in one of these, changing the sample size to compensate). In the last case, the original study used a multi-step procedure where the first step consisted of stimulus generation; we decreased the N in the first stage of the procedure and focused on achieving high power at the second stage.

@@

Effect sizes and Bayes factors improve upon inferences drawn from $p$-values in different ways. The effect size is an estimate of the *magnitude* of an effect, which is stable across sample sizes (assuming a constant experimental procedure, of course) and allows for a more fine-grained comparison across studies than a binary significance criterion. The Bayes Factor is an alternative hypothesis testing method that quantifies the evidence in favor of a hypothesis [@jeffreys1961]. Among other advantages, it does not privilege the null hypothesis and can provide evidence either for or against [see also @scheibehenne2016;@wagenmakers2016].  While the Bayes Factor often agrees with the $p$-value as to which hypothesis is more likely [@wetzels2011], they often substantially disagree on the strength of the evidence.

@@

In this paper, we reported a series of 11 student replications of previously-published experiments from the 2015 volume of *Psychological Science*. We had two interlocking goals: first, to create a proof of concept for pedagogical replications, and second, to provide evidence on the ability of students replicate important, recent findings in a top journal. Importantly, rather than attempting to gauge the truth of particular effects, our project aimed to assess the challenges of replicating findings -- selected on grounds of feasibility and interest -- within the constraints of a course project. Despite this narrower scope, we believe that insights from our approach are applicable to the broader goal of assessing replicability at scale.

>>>>

We eventually used a hybrid, case-by-case selection criterion that attempted to maximize the success of the project while staying within the constraints of our budget.^[We initially were allocated \$1,500 but one student contributed personal research funds, leading to a total cost of $\approx$ \$1,700 for all studies. We attempted to set payment for our studies at approximately \$6/hour based on timing estimated from Pilot A sessions.] For all criteria, experimenters recruited an additional 5% of participants to ensure we still achieved the desired criterion if some participants skipped through or did not complete the experiment. In four cases, we powered the replication attempt to 80\% post-hoc power. In four cases, we used the original sample size. In one case, the original experiment was extremely over-powered in the post-hoc analysis, so we chose a smaller sample size that still would yield high power. In two others, we reduced the number of trials as part of the adaptation to Mechanical Turk (in one of these, changing the sample size to compensate). In the last case, the original study used a multi-step procedure where the first step consisted of stimulus generation; we decreased the N in the first stage of the procedure and focused on achieving high power at the second stage.

@@

Effect sizes and Bayes factors improve upon inferences drawn from $p$-values in different ways. The effect size is an estimate of the *magnitude* of an effect, which is stable across sample sizes (assuming a constant experimental procedure, of course) and allows for a more fine-grained comparison across studies than a binary significance criterion. The Bayes Factor is an alternative hypothesis testing method that quantifies the evidence in favor of a hypothesis [@jeffreys1961]. Among other advantages, it does not privilege the null hypothesis and can provide evidence either for or against a hypothesis [see also @scheibehenne2016;@wagenmakers2016].  While the Bayes Factor often agrees with the $p$-value as to which hypothesis is more likely [@wetzels2011], they often substantially disagree on the strength of the evidence.

@@

In this paper, we reported a series of 11 student replications of previously-published experiments from the 2015 volume of *Psychological Science*. We had two interlocking goals: first, to create a proof of concept for pedagogical replications, and second, to provide evidence on the ability of students to replicate recent findings in a top journal. Importantly, rather than attempting to gauge the truth of particular effects, our project aimed to assess the challenges of replicating findings -- selected on grounds of feasibility and interest -- within the constraints of a course project. Despite this narrower scope, we believe that insights from our approach are applicable to the broader goal of assessing replicability at scale.
